---
title: "Advent of Code 2020 Day 8"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 8: Handheld Halting ---

You have to try to fix the game console of the kid sitting next to you on the flight.

Run your copy of the boot code. Immediately before any instruction is executed a second time, *what value is in the accumulator?*

```{r}

library(tidyverse)

game <- read.delim("data-2020-08.txt", header = FALSE) %>% 
  separate(V1, into = c("inst", "num"), sep = " ") %>% 
  mutate(has_run = FALSE, 
         num = parse_number(num))

```

Create a function for processing and see how it goes

```{r}

run_game <- function(game) {
  game$has_run = FALSE
  x = 1
  acc = 0
  max_x = length(game$inst)

  while(game$has_run[[x]] == FALSE) {
    if (game$inst[[x]] == "acc") {
      acc = acc + game$num[[x]]
      game$has_run[[x]] = TRUE
      x = x + 1
    } else if (game$inst[[x]] == "nop") {
      game$has_run[[x]] = TRUE
      x = x + 1
    } else {
      game$has_run[[x]] = TRUE
      x = x + game$num[[x]]
    }
    if (x > max_x) {
      game_final <<- game
      return(acc)
    } 
  }
  return(NULL)
}

```

Figure things out

```{r}

# maybe it is the last jmp to nop, but we never get that high
# it was great to record where we've been
# what are the possibilities for finishing? 
game <- game %>% 
  mutate(max_pos = num + row_number())
# check out that data
arrange(game, desc(max_pos))
# there's a jmp at 638-170=468
# how do we get there? 
# there are two jumps that equal 467
# so now we've got divergence and we could keep going but nah
# it seems like we could go backwards but that takes much design

val <- run_game(game)

```

## --- Part Two ---

Exactly one instruction is corrupt, either a `jmp` or `nop`, `acc` are all ok.

Fix the program so that it terminates normally by changing exactly one `jmp` (to `nop`) or `nop` (to `jmp`). *What is the value of the accumulator after the program terminates?*

```{r}

for (z in seq_along(game$inst)) {
  if (game$inst[[z]] == "jmp") {
    game$inst[[z]] = "nop"
    revert = "jmp"
  } else if (game$inst[[z]] == "nop") { 
    game$inst[[z]] = "jmp"
    revert = "nop"
  } else {
    next
  }
  val = run_game(game)
  if (is.null(val)) {
    game$inst[[z]] = revert
  } else {
    print(z)
    break
  }
}

val
```
