---
title: "Advent of Code 2020 Day 6"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 6: Custom Customs ---

Customs forms are being handed out on the flight and you need to work out who answers yes to how many questions.

For each group, count the number of questions to which anyone answered "yes". What is the sum of those counts?

```{r}

library(tidyverse)

# read all at once
answers <- read_file("data-2020-06.txt")

```

Unique letters, by group

```{r}

# split into elements with double end line
answer_groups <- str_split(answers, "\r\n\r\n")
# names
names(answer_groups) <- c("ans")
# combo each piece
answer_join <- str_remove_all(answer_groups$ans, "\r\n")

# count letters
counters <- function(lett) {
  sum(str_count(lett, letters) > 0)
}


# apply
count_answers <- map(answer_join, counters)
# reduce
Reduce(`+`, count_answers)

```

## --- Part Two ---

We don't have to work out *any* person answered yes, we have to count *all* who answered yes.

For each group, count the number of questions to which *everyone* answered "yes". *What is the sum of those counts?*

```{r}

count_groups <- function(lett, num) {
  sum(str_count(lett, letters) == num)
}

# part two
answer_split <- str_split(answer_groups$ans, "\r\n")

ans_tbl_rev <- tibble(
                      long_str = answer_join, 
                      ans_piece = answer_split
                     ) %>% 
              mutate(grp_len = lengths(ans_piece), 
                     grp_cnt = map2_dbl(long_str, grp_len, count_groups))
  
Reduce(`+`, ans_tbl_rev$grp_cnt)

```

Tibble calculations

```{r}
ans_tbl <- tibble(ans = answer_split) %>% 
  mutate(grp = row_number()) %>% 
  unnest_longer(ans) %>% 
  mutate(cnt = map(ans, str_count, letters)) %>% 
  nest(data = c("ans", "cnt"))

ans_tbl$num <- 0

for (n in seq_len(nrow(ans_tbl))) {
  tt = ans_tbl$data[[n]]$cnt
  x = tt[[1]]
  if (length(tt) > 1) {
    for (m in seq(2, length(tt))) {
      y = tt[[m]]
      x = bitwAnd(x, y)
    }  
  }
  ans_tbl$num[[n]] = sum(x)
}

# part two answer
sum(ans_tbl$num)
```

Another strategy for part two

```{r}
# part two type 2
split_ans <- function(aa) {
  ans_collect = NULL
  for (i in seq_len(length(aa))) {
    ans_each = str_split(aa[[i]], pattern = "")
    # print(ans_each[[1]])
    if (i == 1) {
      ans_collect = ans_each[[1]]
    } else {
      ans_collect = intersect(ans_collect, ans_each[[1]])
    }
  }
  
  return(glue::glue_collapse(ans_collect))
}

ans_tbl <- tibble(ans = answer_split) %>% 
  mutate(ans_v = map(ans, split_ans), 
         cnt = map(ans_v, counters))


Reduce(`+`, ans_tbl$cnt)
```
