---
title: "Advent of Code 2020 Day 10"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 10: Adapter Array ---

Jolts, and stuff?

Find a chain that uses all of your adapters to connect the charging outlet to your device's built-in adapter and count the joltage differences between the charging outlet, the adapters, and your device. What is the number of 1-jolt differences multiplied by the number of 3-jolt differences?

```{r}

# read the things
jolts <- readLines("data-2020-10.txt")

jolts = c(0, sort(as.numeric(jolts)))

jolts = c(jolts, max(jolts)+3)

```

Sort and then add up the differences

```{r}
ones = 0
threes = 0

for (j in 2:length(jolts)) {
    diff = jolts[j] - jolts[j-1]
    if (diff == 1) {
        ones = ones + 1
    } else {
        if (diff == 3) {
            threes = threes + 1
        }
    }
}

ones * threes

```

## --- Part Two ---

We need to join all of these together.

*What is the total number of distinct ways you can arrange the adapters to connect the charging outlet to your device?*

```{r}

# to connect 0 to 1, is 1 way (and 1 for 0 because I did this manually)
ways = 1
ways_1 = 1
ways_2 = 1
ways_3 = 0

# to connect 1 to 2, or 0 to 2, is 2 ways
ways = 2
ways_1 = 1
ways_2 = 1
ways_3 = 1

# now do the rest for 3 etc
for (j in 4:length(jolts)) {
    # print(c(jolts[j], jolts[j-1], jolts[j-2], jolts[j-3]))
    # print(c(ways, ways_1, ways_2, ways_3))
    if ((jolts[j] - jolts[j-3]) <= 3) {
        ways = ways + ways_3
    }
    if ((jolts[j] - jolts[j-2]) <= 3) {
        ways = ways + ways_2
    }    
    # if ((jolts[j] - jolts[j-1]) <= 3) {
    #     ways = ways + ways_1
    # }
    ways_3 = ways_2
    ways_2 = ways_1
    ways_1 = ways
    # print(ways)
}

print(ways, digits=16)


```
