---
title: "Advent of Code 2020 Day 9"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 9: Encoding Error ---

The data appears to be encrypted with the eXchange-Masking Addition System (XMAS) which, conveniently for you, is an old cypher with an important weakness.

The first step of attacking the weakness in the XMAS data is to find the first number in the list (after the preamble) which is *not* the sum of two of the 25 numbers before it. *What is the first number that does not have this property?*

```{r}

# read the things
code <- read.delim("data-2020-09.txt", header = FALSE)

# setup
tgt_num <- 26
len <- nrow(code)

```

Loop to see what doesn't work

```{r}

for (x in tgt_num:len) {
  found = FALSE
  tgt_sum = code$V1[[x]]
  for (j in (x-25):(x-2)) {
    for (k in (x-24):(x-1)) {
      if (code$V1[[j]] + code$V1[[k]] == tgt_sum) {
        found = TRUE
      }
      if (found) break
    }
    if (found) break
  }
  if (!found) {
    print(tgt_sum)
    print(x)
    break
  }
}

```

## --- Part Two ---

There is a contiguous sequence of numbers that add up to the missing number. Find them!

*What is the encryption weakness in your XMAS-encrypted list of numbers?*

```{r}


tgt <- 1639024365
tgt_loc <- 654

found <- FALSE
# part two
for (qq in 3:50) { 
  for (x in 1:tgt_loc) {
    chk_sum = sum(code$V1[x:(x+qq)])
    if (chk_sum == tgt) {
      chk_vect = code$V1[x:(x+qq)]
      part_two = sum(min(chk_vect), max(chk_vect))
      print(x)
      print(qq)
      print(paste("part two answer:", part_two))
      found = TRUE
    }
    if (found) break
  }
  if (found) break
}


```
