---
title: "2021 Day 3"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 3: Binary Diagnostic ---

Use the binary numbers in your diagnostic report to calculate the gamma rate and epsilon rate, then multiply them together. *What is the power consumption of the submarine?* (Be sure to represent your answer in decimal, not binary.)

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-03.txt")
x <- readLines(my_file)

```

Maybe write an explanation of the solution approach here

```{r}

# split into individual columns
bins = str_split_fixed(x, "", 12) %>% as.data.frame()

# need them to be numeric
bin_num = sapply(bins, as.numeric)

gamma_rate =  strtoi(paste(as.numeric(colSums(bin_num) >= 500), collapse = ""), base = 2)

epsilon_rate = strtoi(paste(as.numeric(colSums(bin_num) < 500), collapse = ""), base = 2)

part_one = gamma_rate * epsilon_rate
  
part_one
```

## --- Part Two ---

Use the binary numbers in your diagnostic report to calculate the oxygen generator rating and CO2 scrubber rating, then multiply them together. *What is the life support rating of the submarine?* (Be sure to represent your answer in decimal, not binary.)

```{r}

return_oxygen <- function(z, colnum = 1) {
  # z is a matrix of binary digits
  # calculate the length
  # then which digit is more common in the column
  # input of colnum

  # number of rows of array
  lngth = nrow(z)

  # wdth for stopping
  wdth = ncol(z)

  # column with greater number of digits
  dgt = as.integer(sum(z[,colnum]) >= ceiling(lngth / 2))

  # filter (is that the right word for non-tidyverse?)
  row_bool = z[,colnum] == dgt
  z = z[row_bool, , drop = FALSE]

  # return if we are at the end
  if (colnum == wdth) {
    print("colnum = 12")
    return(z)
  }

  if (nrow(z) == 1) {
    print("only one row")
    return(z)
  }

  # increment by 1
  colnum = colnum + 1

  # call same function again with new column to check
  z = return_oxygen(z, colnum)

}

return_c02 <- function(z, colnum = 1) {
  # this is the same as above but with less than sign

  # number of rows of array
  lngth = nrow(z)

  # wdth for stopping
  wdth = ncol(z)

  # column with greater number of digits
  dgt = as.integer(sum(z[,colnum]) < ceiling(lngth / 2))

  # filter (is that the right word for non-tidyverse?)
  row_bool = z[,colnum] == dgt
  z = z[row_bool, , drop = FALSE]

  # return if we are at the end
  if (colnum == wdth) return(z)

  if (nrow(z) == 1) return(z)

  # increment by 1
  colnum = colnum + 1

  # call same function again with new column to check
  z = return_c02(z, colnum)

}


```

And then use the functions

```{r}

# width of one string
wdth = nchar(x[[1]])

# split into individual columns
bins = str_split_fixed(x, "", wdth) %>% as.data.frame()

# need them to be numeric
bin_num = sapply(bins, as.numeric)

# recursive function call
oxygen_binary = return_oxygen(bin_num)

# convert to integer for oxygen
oxygen_rating = strtoi(paste(oxygen_binary, collapse = ""), base = 2)

# recursive with less than, c02
c02_binary = return_c02(bin_num)

# convert to integer
c02_rating = strtoi(paste(c02_binary, collapse = ""), base = 2)

# answer
oxygen_rating * c02_rating
  
```
