---
title: "2021 Day 19"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 19: Beacon Scanner ---

As your [probe](https://adventofcode.com/2021/day/17) drifted down through this area, it released an assortment of *beacons* and *scanners* into the water. It's difficult to navigate in the pitch black open waters of the ocean trench, but if you can build a map of the trench using data from the scanners, you should be able to safely reach the bottom.

Assemble the full map of beacons. *How many beacons are there?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-19.txt")
x <- readLines(my_file)

```

I think the way to go with this is the distance between beacons? That will make a big matrix. Maybe the angle between beacons will have to be involved too.

```{r}

bcn = tibble(zz = x) %>%
    separate(zz, into = c("x", "y", "z"), sep = ",", remove = FALSE) %>%
    mutate(scn = ifelse(str_detect(zz, "scanner"),
                         parse_number(substr(zz, 5, 15)),
                         NA)) %>%
    fill(scn, .direction = "down") %>%
    filter(!is.na(z)) %>%
    select(-zz) %>%
    group_by(scn) %>%
    mutate(bb = row_number()) %>%
    mutate(across(c(x, y, z), as.numeric)) %>%
    ungroup()

bcn_cross = full_join(bcn, bcn, 
                      by = c("scn"), 
                      suffix = c(".l", ".r"),
                      relationship = "many-to-many") %>%
    filter(bb.l < bb.r) %>%
    mutate(dd = round(sqrt((x.l - x.r)**2 + (y.l - y.r)**2 + (z.l - z.r)**2), 2))

dist_order = bcn_cross %>% arrange(dd)

distinct_distance <- bcn_cross %>% distinct(dd)

beacon_table <- table(bcn_cross$dd) |> 
    as_tibble(.name_repair="unique") |> 
    filter(n>1)


```

Paste part two here
