---
title: "2021 Day 8"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 8: Seven Segment Search ---

You need to find your way through a cave with some malfunctioning digital number displays.

*In the output values, how many times do digits `1`, `4`, `7`, or `8` appear?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-08.txt")
x <- readLines(my_file)

```

These digits are ones that have the right hand side line.

*Author's note: was I really good at R or did I copy this from the R for Data Science slack chat?*

```{r}

# split pieces by pipe
# pipe is a character delimiter, need to escape
segments = tibble(parts = x) %>%
    separate(parts, into = c("nums", "digi"), sep = " \\| ", remove = FALSE)

poss = c(2,3,4,7)

digits = segments %>%
    separate(digi, into = c("d1", "d2", "d3", "d4"), remove = FALSE) %>%
    mutate(across(d1:d4, ~ nchar(.x) %in% poss))

sum(digits %>% summarise(across(d1:d4, sum)))
```

## --- Part Two ---

For each entry, determine all of the wire/segment connections and decode the four-digit output values. *What do you get if you add up all of the output values?*

Functions!

```{r}

get_nums <- function(vct, cnt) {
  paste0(sort(unlist(strsplit(vct[nchar(vct) == cnt], ""))), collapse = "")
}

get_cnt <- function(vct, cnt) {
  letters[1:7][str_count(paste0(vct, collapse = ""), letters[1:7]) == cnt]
}

get_a <- function(v1, v2) {
  setdiff(unlist(strsplit(v1, "")), unlist(strsplit(v2, "")))
}

get_d <- function(four, one, b) {
  setdiff(setdiff(unlist(strsplit(four, "")), unlist(strsplit(one, ""))), b)
}

get_g <- function(one, seven, four, e) {
  mid = union(union(unlist(strsplit(one, "")), unlist(strsplit(seven, ""))), unlist(strsplit(four, "")))
  setdiff(setdiff(letters[1:7], mid), e)
}

get_c <- function(nn, aa) {
  dub = get_cnt(nn, 8)
  setdiff(dub, aa)
}

do_six <- function(z1, z2, z3, z4, z5, z6) {
  paste0(sort(c(z1, z2, z3, z4, z5, z6)), collapse = "")
}

do_five <- function(z1, z2, z3, z4, z5) {
  paste0(sort(c(z1, z2, z3, z4, z5)), collapse = "")
}

get_eight <- function(xx) {
  paste0(letters[1:7], collapse = "")
}

get_digi <- function(dd) {
  lapply(strsplit(dd, ""), sort)
}

```

Use the functions and create many columns for possibilities

```{r}

# so much mutate
segments = tibble(parts = x) %>%
    separate(parts, into = c("nums", "digi"), sep = " \\| ") %>%
    mutate(nums = str_split(nums, " "),
           n_1 = map(nums, get_nums, 2),
           n_7 = map(nums, get_nums, 3),
           n_4 = map(nums, get_nums, 4),
           e_ = map(nums, get_cnt, 4),
           b_ = map(nums, get_cnt, 6),
           f_ = map(nums, get_cnt, 9),
           a_ = map2(n_7, n_4, get_a),
           d_ = pmap(list(n_4, n_1, b_), get_d),
           g_ = pmap(list(n_1, n_7, n_4, e_), get_g),
           c_ = map2(nums, a_, get_c),
           n_0 = pmap(list(a_, b_, c_, e_, f_, g_), do_six),
           n_6 = pmap(list(a_, b_, d_, e_, f_, g_), do_six),
           n_9 = pmap(list(a_, b_, c_, d_, f_, g_), do_six),
           n_2 = pmap(list(a_, c_, d_, e_, g_), do_five),
           n_3 = pmap(list(a_, c_, d_, f_, g_), do_five),
           n_5 = pmap(list(a_, b_, d_, f_, g_), do_five),
           n_8 = map(8, get_eight),
           rr = row_number()
          ) %>%
    pivot_longer(starts_with("n_")) %>%
    rowwise() %>%
    mutate(value = paste0(value, collapse = "")) %>%
    select(!ends_with("_"), -digi, -nums)


digits = tibble(parts = x) %>%
    separate(parts, into = c("nums", "digi"), sep = " \\| ") %>%
    mutate(digi = str_split(digi, " "),
           digis = map(digi, get_digi),
           rr = row_number()) %>%
    unnest_wider(digis, names_sep = "_") %>%
    pivot_longer(starts_with("digis_")) %>%
    rowwise() %>%
    mutate(value = paste0(value, collapse = "")) %>%
    select(-nums, -digi)

results = inner_join(digits, segments, by = c("rr", "value")) %>%
    mutate(nums = parse_number(name.y)) %>%
    pivot_wider(id_cols = rr, names_from = name.x, values_from = nums) %>%
    rowwise() %>%
    mutate(val = as.integer(paste0(c(digis_1, digis_2, digis_3, digis_4), collapse = "")))

sum(results$val)
```
