---
title: "2021 Day 14"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 14: Extended Polymerization ---

We need to reinforce the submarine with some polymers.

Apply 10 steps of pair insertion to the polymer template and find the most and least common elements in the result. *What do you get if you take the quantity of the most common element and subtract the quantity of the least common element?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-14.txt")
x <- readLines(my_file)

```

Split the input to appropriate pieces.

```{r}

pol = unlist(strsplit(x[[1]], ""))

rules = strsplit(x[3:length(x)], " -> ")

rr = list()

for (r in 1:length(rules)) {
    rr[rules[[r]][1]] = rules[[r]][2]
}

```

Loop through ten times.

```{r}

old_list = pol
new_list = vector()

for (i in 1:10) {
    # figure out the length to loop
    loop = length(old_list) - 1
    for (j in 1:loop) {
        # allocate elements to new list
        new_list = c(new_list, old_list[j])
        # get the substitute key
        key = paste0(old_list[j], old_list[j+1])
        if (key %in% names(rr)){
            # print(rr[key])
            new_list = c(new_list, unlist(rr[key]))
        } else {
            print(key)
        }
    }
    new_list = c(new_list, old_list[j+1])
    old_list = new_list
    new_list = vector()
}

sort(table(old_list))

3108 - 506

```

## --- Part Two ---

Apply *40* steps of pair insertion to the polymer template and find the most and least common elements in the result. *What do you get if you take the quantity of the most common element and subtract the quantity of the least common element?*

```{r}

# there might be a pattern to this
# because 40 doesn't sound that big
# but it is probably massively long lists? 
# or maybe I need to only keep track of how many times characters appear
# brute force to start
# 20 is too much to start

# old_list = pol
# new_list = vector()
# 
# for (i in 1:20) {
#     # figure out the length to loop
#     loop = length(old_list) - 1
#     for (j in 1:loop) {
#         # allocate elements to new list
#         new_list = c(new_list, old_list[j])
#         # get the substitute key
#         key = paste0(old_list[j], old_list[j+1])
#         if (key %in% names(rr)){
#             # print(rr[key])
#             new_list = c(new_list, unlist(rr[key]))
#         } else {
#             print(key)
#         }
#     }
#     new_list = c(new_list, old_list[j+1])
#     old_list = new_list
#     new_list = vector()
# }
# 
# sort(table(old_list))

```

Another approach: go through and see if there are consistent amounts of letters added per round.

```{r}

for (k in 1:8) {
    old_list = pol
    new_list = vector()
    
    for (i in 1:k) {
        # figure out the length to loop
        loop = length(old_list) - 1
        for (j in 1:loop) {
            # allocate elements to new list
            new_list = c(new_list, old_list[j])
            # get the substitute key
            key = paste0(old_list[j], old_list[j+1])
            if (key %in% names(rr)){
                # print(rr[key])
                new_list = c(new_list, unlist(rr[key]))
            } else {
                print(key)
            }
        }
        new_list = c(new_list, old_list[j+1])
        old_list = new_list
        new_list = vector()
    }
    
    print(sort(table(old_list)))
}
```
