---
title: "2021 Day 11"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 11: Dumbo Octopus ---

Watch our for the bioluminescent octopuses.

Given the starting energy levels of the dumbo octopuses in your cavern, simulate 100 steps. *How many total flashes are there after 100 steps?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-11.txt")
x <- readLines(my_file)

```

Function!

```{r}

increment_octo <- function(arr, vct) {
  for (j in seq(2, 11)) {
    for (k in seq(2, 11)) {
      pos = (k - 1) * 12 + j
      if (arr[j, k] > 9 & pos %in% vct) {
        # print(pos)
        arr[(j-1):(j+1), (k-1):(k+1)] = arr[(j-1):(j+1), (k-1):(k+1)] + 1
      }
    }
  }
  return(arr)
}

```

Loops!

```{r}

# array, I think
octo_s = stringr::str_split_fixed(x, "", 10) %>% as.data.frame()
octo_s = sapply(octo_s, as.numeric)

octo = array(NA, c(12, 12))
octo[2:11, 2:11] = octo_s
# keep track of how many flashes there have been
counter = 0

# loop through 100 times
# is part two going to be a million times?
for (j in seq(100)) {
    # print(paste("iteration: ", j))
    # add one to the array
    octo = octo + 1
    # are any octopi going to flash?
    flash = which(octo > 9)
    # if none flash, go to next iteration
    if (!any(flash, na.rm = TRUE)) next

    # keep track of which have already flashed
    new_flash = vector()
    old_flash = vector()
    # if we do have some flashing
    repeat {
        # new flash, iterate these ones
        new_flash = setdiff(flash, old_flash)
        # allocate to keep track
        old_flash = flash
        # this function takes care of iteration
        octo = increment_octo(octo, new_flash)
        # get a new list of flashers
        flash = which(octo > 9)
        # break the repeat if we aren't flashing any more
        # length might not be the best test but we can only add so it's ok?
        if (length(flash) == length(old_flash)) break
    }
    octo[octo > 9] = 0
    counter = counter + length(flash)
    # print(paste("length of flashes: ", length(flash)))

}

counter 
```

## --- Part Two ---

If you can calculate the exact moments when the octopuses will all flash simultaneously, you should be able to navigate through the cavern. *What is the first step during which all octopuses flash?*

```{r}

# array, I think
octo_s = stringr::str_split_fixed(x, "", 10) %>% as.data.frame()
octo_s = sapply(octo_s, as.numeric)

octo = array(NA, c(12, 12))
octo[2:11, 2:11] = octo_s
# keep track of how many flashes there have been
counter = 0

# loop through 100 times
# is part two going to be a million times?
for (j in seq(1000)) {
    # print(paste("iteration: ", j))
    # add one to the array
    octo = octo + 1
    # are any octopi going to flash?
    flash = which(octo > 9)
    # if none flash, go to next iteration
    if (!any(flash, na.rm = TRUE)) next

    # keep track of which have already flashed
    new_flash = vector()
    old_flash = vector()
    # if we do have some flashing
    repeat {
        # new flash, iterate these ones
        new_flash = setdiff(flash, old_flash)
        # allocate to keep track
        old_flash = flash
        # this function takes care of iteration
        octo = increment_octo(octo, new_flash)
        # get a new list of flashers
        flash = which(octo > 9)
        # break the repeat if we aren't flashing any more
        # length might not be the best test but we can only add so it's ok?
        if (length(flash) == length(old_flash)) break
    }
    octo[octo > 9] = 0
    counter = counter + length(flash)
    # print(paste("length of flashes: ", length(flash)))
    if (length(flash) == 100) break
}

counter

```
