---
title: "2021 Day 10"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 10: Syntax Scoring ---

The navigation subsystem is corrupted! There are so many brackets.

Find the first illegal character in each corrupted line of the navigation subsystem. *What is the total syntax error score for those errors?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-10.txt")
x <- readLines(my_file)

```

Write a function to get the error bracket

```{r}

return_error <- function(x) {
  new = x
  pairs = c("\\[\\]", "\\(\\)", "\\{\\}", "\\<\\>")
  backs = "[\\)\\}\\]\\>]"

  repeat {
    old = new
    for (j in seq_along(pairs)) {
      new = str_replace_all(new, pairs[j], "")
    }
    if (old == new) break
  }

  str_match(new, backs)[[1]]
}

```

Simple script to loop through the inputs and assign scores

```{r}

# replace the inner matches until we can replace no more
pieces = tibble(parts = x) %>%
    rowwise() %>%
    mutate(closer = return_error(parts))

points = tribble(
    ~closer, ~points,
    ")", 3,
    "]", 57,
    "}", 1197,
    ">", 25137,
  )

part_one = inner_join(pieces, points, by='closer')
sum(part_one$points)

```

## --- Part Two ---

Find the completion string for each incomplete line, score the completion strings, and sort the scores. *What is the middle score?*

```{r}

return_closers <- function(x) {
  new = x
  pairs = c("\\[\\]", "\\(\\)", "\\{\\}", "\\<\\>")
  openers = c("\\[", "\\(", "\\{", "\\<")
  closers = c("\\]", "\\)", "\\}", "\\>")
  backs = "[\\)\\}\\]\\>]"

  repeat {
    old = new
    for (j in seq_along(pairs)) {
      new = str_replace_all(new, pairs[j], "")
    }
    if (old == new) break
  }

  if (!is.na(str_match(new, backs)[[1]])) return("")

  forward = stringi::stri_reverse(new)

  backward = forward
  for (k in seq_along(openers)) {
    backward = str_replace_all(backward, openers[k], closers[k])
  }
  return(backward)
}

score_seq <- function(z) {
  pp = unlist(strsplit(z, ""))

  cc = ")]}>"

  ss = 0
  for (j in seq_along(pp)) {
    ss = ss * 5
    ss = ss + str_locate(cc, paste0("\\", pp[[j]]))[[1]]
  }
  return(ss)
}
```

Again, create functions and then use those for a nice tibble

```{r}

pieces = tibble(parts = x) %>%
    rowwise() %>%
    mutate(closer = return_closers(parts),
           scores = score_seq(closer)) %>%
    filter(scores != 0)

scores = sort(pieces$scores)
scores[ceiling(length(scores) / 2)]

```
