---
title: "2021 Day 13"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 13: Transparent Origami ---

```         
Congratulations on your purchase! To activate this infrared thermal imaging camera system, please enter the code found on page 1 of the manual.
```

The manual is transparent paper with dots and folding instructions.

*How many dots are visible after completing just the first fold instruction on your transparent paper?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-13.txt")
x <- readLines(my_file)

```

Maybe write an explanation of the solution approach here

```{r}

inputs <- tibble(inpt = x) %>%
    separate(inpt, c("X", "Y"), sep = ",", convert = TRUE)

folds = inputs %>%
    filter(is.na(Y) & X != '') %>%
    select(-Y) %>%
    separate(X, c("dir", "num"), sep = "=", convert = TRUE) %>%
    mutate(dir = str_remove(dir, "fold along "))

thermal = inputs %>%
    filter(!is.na(Y)) %>%
    mutate(X = as.numeric(X))

for (j in seq_along(folds)) {
    if (folds$dir[[j]] == "x") {
        for (k in seq_along(thermal$X)) {
            if (thermal$X[[k]] > folds$num[[j]]){
                thermal$X[[k]] = folds$num[[j]] - (thermal$X[[k]] - folds$num[[j]])
            }
        }
    } else {
        for (k in seq_along(thermal$Y)) {
            if (thermal$Y[[k]] > folds$num[[j]]){
                thermal$Y[[k]] = folds$num[[j]] - (thermal$Y[[k]] - folds$num[[j]])
            }
        }
    }
    break
}

nrow(thermal %>% distinct())
```

## --- Part Two ---

Finish folding the transparent paper according to the instructions. The manual says the code is always *eight capital letters*.

*What code do you use to activate the infrared thermal imaging camera system?*

```{r}

inputs <- tibble(inpt = x) %>%
    separate(inpt, c("X", "Y"), sep = ",", convert = TRUE)

folds = inputs %>%
    filter(is.na(Y) & X != '') %>%
    select(-Y) %>%
    separate(X, c("dir", "num"), sep = "=", convert = TRUE) %>%
    mutate(dir = str_remove(dir, "fold along "))

thermal = inputs %>%
    filter(!is.na(Y)) %>%
    mutate(X = as.numeric(X))

for (j in seq_along(folds$dir)) {
    if (folds$dir[[j]] == "x") {
        for (k in seq_along(thermal$X)) {
            if (thermal$X[[k]] > folds$num[[j]]){
                thermal$X[[k]] = folds$num[[j]] - (thermal$X[[k]] - folds$num[[j]])
            }
        }
    } else {
        for (k in seq_along(thermal$Y)) {
            if (thermal$Y[[k]] > folds$num[[j]]){
                thermal$Y[[k]] = folds$num[[j]] - (thermal$Y[[k]] - folds$num[[j]])
            }
        }
    }
}

arr = array(".", c(max(thermal$Y)+1, max(thermal$X)+1))

for (k in seq_along(thermal$X)) {
    arr[thermal$Y[[k]]+1, thermal$X[[k]]+1] = '#'
}

for (a in 1:6) { 
    print(paste(arr[a,], collapse=""))
}

```
