---
title: "2021 Day 2"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 2: Dive! ---

We have to figure out where the submarine is going.

Calculate the horizontal position and depth you would have after following the planned course. *What do you get if you multiply your final horizontal position by your final depth?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-02.txt")
x <- readLines(my_file)

```

up and down mean different things, but we can handle that.

```{r}

# create a helper object to provide instructions
effect <- tribble(
~direction, ~x, ~depth,
"forward", 1, 0,
"down", 0, 1,
"up", 0, -1
)

# create direction and value in their own columns
instruct <- tibble(inpt = x) %>%
separate(inpt, c("direction", "val"), convert = TRUE)

# join and calculate change at each step
delta_sub <- inner_join(instruct, effect, by = c("direction")) %>%
mutate(x_delta = val * x,
       depth_delta = val * depth)

# part one answer is sum of final x and depth position
sum(delta_sub$x_delta) * sum(delta_sub$depth_delta)

```

## --- Part Two ---

Actually, the instructions are different.

Using this new interpretation of the commands, calculate the horizontal position and depth you would have after following the planned course. *What do you get if you multiply your final horizontal position by your final depth?*

```{r}

# create a helper object to provide instructions
effect <- tribble(
~direction, ~x, ~dp,
"forward", 1, 0,
"down", 0, 1,
"up", 0, -1
)

# create direction and value in their own columns
instruct <- tibble(inpt = x) %>%
separate(inpt, c("direction", "val"), convert = TRUE)

# join and calculate change at each step
delta_sub <- inner_join(instruct, effect, by = c("direction")) %>%
mutate(x_delta = val * x,
       depth_delta = val * dp,
       aim = cumsum(depth_delta),
       depth_aim = aim * x_delta)

# part one answer is sum of final x and depth position
sum(delta_sub$x_delta) * sum(delta_sub$depth_aim)

```
