---
title: "2021 Day 17"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 17: Trick Shot ---

Send a probe into the ocean trench.

Find the initial velocity that causes the probe to reach the highest `y` position and still eventually be within the target area after any step. *What is the highest `y` position it reaches on this trajectory?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-17.txt")
x <- readLines(my_file)

nums = str_extract_all(x, '-?\\d+')

x0 = as.integer(nums[[1]][1])
x1 = as.integer(nums[[1]][2])
y0 = as.integer(nums[[1]][3])
y1 = as.integer(nums[[1]][4])

```

We can probably try out all options and figure this out. Part two guess: the target area is multiplied by 100 or 1,000,000 and brute force becomes impossible.

```{r}

# make a function so we can reuse program

cannonball <- function(v) {
    pos = c(0,0)
    max_y = -999
    for (i in 1:1000) {
        # move cannonball
        pos[1] = pos[1] + v[1]
        pos[2] = pos[2] + v[2]
        if (pos[2] > max_y) {
            max_y = pos[2]
        }
        # print for debug
        # print(i)
        # print(pos)
        # print(v)
        # test position
        if (pos[1] >= x0 && pos[1] <= x1) { 
            if (pos[2] >= y0 && pos[2] <= y1) {
                # found a good spot
                break
            }    
        }
        # edit velocity
        # x must always be positive for my target, I think
        if (v[1] > 0) {
            v[1] = v[1] - 1
        } else {
            v[1] = 0
        }
        v[2] = v[2] - 1
    }
    return(c(i, max_y))
}



```

Use this function and see what we can find

```{r}

# cannonball(c(14,7))

# has to be larger than 12, otherwise we don't get into the square
# 13: 5, 15; 6, 21; 
# 14: 6, 21; 7, 28; 

# Loop!
for (j in 10:30) {
    max_y = 0
    step = 0
    for (k in 5:1000) { 
        res = cannonball(c(j, k))
        # print(res)
        if (res[1] < 1000) { 
            if (res[2] > max_y) { 
                max_y = res[2]
                step = res[1]
            }
        }
    }
    print(c(j, step, max_y))
}

```

## --- Part Two ---

*How many distinct initial velocity values cause the probe to be within the target area after any step?*

```{r}
all_count = 0

# Loop!
# Negatives!
for (j in 10:140) {
    counter = 0
    for (k in -200:1000) { 
        res = cannonball(c(j, k))
        # print(res)
        if (res[1] < 1000) { 
            counter = counter + 1
        }
    }
    print(c(j, counter))
    all_count = all_count + counter
}

print(all_count)

```
