---
title: "2021 Day 25"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 25: Sea Cucumber ---

The floor is just there, but it is covered in sea cucumbers.

Find somewhere safe to land your submarine. *What is the first step on which no sea cucumbers move?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-25.txt")
x <- readLines(my_file)

```

Loop a little to make sure we are moving properly, and then let the cucumbers move.

```{r}

# prepare an array
cuc_width = nchar(x[[1]])
cuc_length = length(x)

cuc = stringr::str_split_fixed(x, "", cuc_width)

```

Test and move

```{r}

counter = 0

while (counter < 1) {
    movement = 0
    counter = counter + 1
    
    # test the easterly cucumbers: >
    for (i in seq(1,cuc_length)) {
        for (j in seq(1,cuc_width)) {
            if (cuc[i,j] == ">") {
                if (j == cuc_width) {
                    # if we're at the end of the row, check back
                    if (cuc[i,1] == ".") {
                        cuc[i,j] = "."
                        cuc[i,1] = ">"
                        movement = movement + 1      
                        
                    }
                } else {
                    # check the next cell
                    if (cuc[i,j+1] == ".") {
                        cuc[i,j] = "."
                        cuc[i,j+1] = ">"
                        movement = movement + 1                        
                    }                      
                }
             
            }
        }
    }        

    # test the southerly cucumbers: v 
    for (j in seq(1,cuc_width)) {
        for (i in seq(1,cuc_length)) {
#    for (i in seq(1,cuc_length)) {
#        for (j in seq(1,cuc_width)) {
            if (cuc[i,j] == "v") {
                if (i == cuc_length) {
                    # we're at the bottom, check the top
                    if (cuc[1,j] == ".") {
                        cuc[i,j] = "."
                        cuc[1,j] = "v"
                        movement = movement + 1                        
                    }
                } else {
                    # test for space
                    if (cuc[i+1,j] == ".") {
                        cuc[i,j] = "."
                        cuc[i+1,j] = "v"
                        movement = movement + 1                        
                    }                    
                }
            }

        }
    }   
    
    if (movement == 0) { 
        break
    }
    # print(movement)
}
    
# 462 too high
# 448 too high 

```

The above catches the cucumbers that have already moved and I don't think there's an iteration schema that can move without collecting them, unless I iterate backwards? But that let's them move if they should be blocked.

So, `which` for an efficient loop of where they are, collecting possible moves.

```{r}

counter = 0

while (counter < 1000) {

    # Easterly Cucumbers
    east = which(cuc == ">", arr.ind = TRUE)
    east = cbind(east, FALSE)
    # print(nrow(east))
        
    for (e in seq_len(nrow(east))) {
        if (east[e,2] == cuc_width) {
            # check the start of the row
            if (cuc[east[e,1], 1] == ".") {
                east[e,3] = TRUE
            }
        } else if (cuc[east[e,1], east[e,2]+1] == ".") {
            # if we can move, add it to moves
            east[e,3] = TRUE
        }
    }
    
    # move only the ones that we think can move
    east = subset(east, east[,3] == TRUE)
    
    # do moves
    for (m in seq_len(nrow(east))) {
        # do the moves
        if (east[m,2] == cuc_width) {
            cuc[east[m,1], 1] = ">"
            cuc[east[m,1], east[m,2]] = "."
        } else {
            cuc[east[m,1], east[m,2]+1] = ">"
            cuc[east[m,1], east[m,2]] = "."
        }
    }
    
    # Southerly Cucumbers
    south = which(cuc == "v", arr.ind = TRUE)
    south = cbind(south, FALSE)
    # print(nrow(south))
    
    for (s in seq_len(nrow(south))) {
        if (south[s,1] == cuc_length) {
            # check the top of the column
            if (cuc[1, south[s,2]] == ".") {
                south[s,3] = TRUE
            }
        } else if (cuc[south[s,1] + 1, south[s,2]] == ".") {
            # check for moves
            south[s,3] = TRUE
        }
    }
    
    # moves the ones we want to move  
    south = subset(south, south[,3] == TRUE)
    
    for (m in seq_len(nrow(south))) {
        # do the moves
        if (south[m,1] == cuc_length) {
            cuc[1, south[m,2]] = "v"
            cuc[south[m,1], south[m,2]] = "."
        } else {
            cuc[south[m,1] + 1, south[m,2]] = "v"
            cuc[south[m,1], south[m,2]] = "."
        }
    }
    
    if (nrow(east) + nrow(south) == 0) {
        break
    } else {
        # print(nrow(east) + nrow(south))
        counter = counter + 1
    }
    
}

# 360 someone else's anwer
# 361 incorrect
# 362 incorrect

# we don't move on that last run through - one less

```
