---
title: "2021 Day 22"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 22: Reactor Reboot ---

We need to reboot the reactor of the submarine.

Execute the reboot steps. Afterward, considering only cubes in the region `x=-50..50,y=-50..50,z=-50..50`, *how many cubes are on?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-22.txt")
x <- readLines(my_file)

```

Maybe write an explanation of the solution approach here

```{r}

cube = tibble(zz = x) %>%
    separate(zz,
             into = c("dd", "x_", "x1", "b_", "x2", "y_", "y1", 
                      "c_", "y2", "z_", "z1", "d_", "z2"),
             sep = "[, =\\.]") %>%
    select(!ends_with("_")) %>%
    mutate(across(x1:z2, as.integer),
           across(x1:z2, ~ . + 50))

reboot = array(0, c(101, 101, 101))


for (i in seq(20)) {
    if (cube$dd[i] == "on") {
        chng = 1
    } else {
        chng = 0
    }
    reboot[cube$x1[i]:cube$x2[i], 
           cube$y1[i]:cube$y2[i], 
           cube$z1[i]:cube$z2[i]] = chng
}

sum(reboot)

```

## --- Part Two ---

Do all of the steps for all of the cubes!

Starting again with all cubes *off*, execute all reboot steps. Afterward, considering all cubes, *how many cubes are on?*

```{r}

# create instructions
cube = tibble(zz = x) %>%
    separate(zz,
             into = c("dd", "x_", "x1", "b_", "x2", "y_", "y1", 
                      "c_", "y2", "z_", "z1", "d_", "z2"),
             sep = "[, =\\.]") %>%
    select(!ends_with("_")) %>%
    mutate(across(x1:z2, as.integer))

# reboot needs to be smarter than initialising the whole thing
cube_min = cube %>%
    summarise(across(x1:z2, min))
# 
# # add 100,000 to everything
# cube = cube %>% mutate(across(x1:z2, ~ . + 100000))
# 
# # list of arrays
# reboot = vector("list", nrow(cube))
# 
# # loop through instructions
# for (i in seq_len(nrow(cube))) {
#     xx = abs(cube$x1[i] - cube$x2[i]) + 1
#     yy = abs(cube$y1[i] - cube$y2[i]) + 1
#     zz = abs(cube$z1[i] - cube$z2[i]) + 1
# 
#     if (cube$dd[i] == "on") {
#         arr = array(1, c(xx, yy, zz),
#                   list(cube$x1[i]:cube$x2[i], 
#                        cube$y1[i]:cube$y2[i], 
#                        cube$z1[i]:cube$z2[i]))
# 
#       reboot[[i]] = arr
# 
#     } else {
#         chng = 0
#     }
# }
# 
# # answer
# sum(unlist(reboot))


```

```         
Error: cannot allocate vector of size 49416.7 Gb
```
