---
title: "2021 Day 4"
author: "Nathan Moore"
format: html
editor: visual
---

## --- Day 4: Giant Squid ---

Let's play bingo against a giant squid!

To guarantee victory against the giant squid, figure out which board will win first. *What will your final score be if you choose that board?*

```{r}

library(tidyverse)

my_file <- here::here("2021", "data-2021-04.txt")
x <- readLines(my_file)

```

functions

```{r}

check_cards_1 <- function(cards) {
  # group by card and row
  # group by card and col

  row_bingo = cards %>%
    group_by(card_num, row_num) %>%
    summarise(bingo = sum(called), .groups = "drop") %>%
    filter(bingo == 5)

  if (nrow(row_bingo) == 1) return (row_bingo$card_num[[1]])

  col_bingo = cards %>%
    group_by(card_num, col_num) %>%
    summarise(bingo = sum(called), .groups = "drop") %>%
    filter(bingo == 5)

  if (nrow(col_bingo) == 1) return (col_bingo$card_num[[1]])

  return(FALSE)

}


check_rows <- function(cards) {
  # group by card and row

  row_bingo = cards %>%
    filter(bingo_done == FALSE) %>%
    group_by(card_num, row_num) %>%
    summarise(bingo = sum(called), .groups = "drop") %>%
    filter(bingo == 5)

  if (nrow(row_bingo) >= 1) return (row_bingo$card_num)

  return(FALSE)

}


check_cols <- function(cards) {
  # group by card and col

  col_bingo = cards %>%
    filter(bingo_done == FALSE) %>%
    group_by(card_num, col_num) %>%
    summarise(bingo = sum(called), .groups = "drop") %>%
    filter(bingo == 5)

  if (nrow(col_bingo) >= 1) return (col_bingo$card_num)

  return(FALSE)

}


create_long_card_record <- function(z) {
  # separate the columns
  # number the bingo cards
  # number the rows
  # remove na
  tibble(bb = z) %>%
    separate(bb, c("c1", "c2", "c3", "c4", "c5"), sep = c(3, 6, 9, 12), convert = TRUE) %>%
    mutate(row_na = is.na(c1),
           card_num = cumsum(row_na)) %>%
    group_by(card_num) %>%
    mutate(row_num = row_number() - 1) %>%
    ungroup() %>%
    filter(row_num != 0) %>%
    select(-row_na) %>%
    pivot_longer(c("c1", "c2", "c3", "c4", "c5"), names_to = "col_num") %>%
    mutate(col_num = parse_number(col_num),
           called = FALSE)
}

```

We get the numbers and the cards, and figure out which cards are good. Score them all.

```{r}

# bingo numbers, first row in text file
bingo_numbers = as.numeric(unlist(strsplit(x[1], ",")))

# bingo cards
bingo_cards = create_long_card_record(x[2:length(x)])

# "call" our bingo numbers but stop if we have a bingo
for (j in 1:length(bingo_numbers)) {
    my_num = bingo_numbers[j]
    # print(my_num)
    bingo_cards = bingo_cards %>%
      mutate(called = (called | value == my_num))
    
    do_we_have_bingo = check_cards_1(bingo_cards)
    
    if (do_we_have_bingo) break

}

# winning card numbers that haven't been called yet
uncalled = bingo_cards %>%
    filter(called == FALSE & card_num == do_we_have_bingo) %>%
    summarise(win_card = sum(value))

# part one answer
uncalled$win_card[[1]] * my_num


```

## --- Part Two ---

Maybe we should let the squid win.

Figure out which board will win last. *Once it wins, what would its final score be?*

```{r}

# bingo numbers, first row in text file
bingo_numbers = as.numeric(unlist(strsplit(x[1], ",")))

# bingo cards
bingo_cards = create_long_card_record(x[2:length(x)])

bingo_cards = bingo_cards %>%
mutate(bingo_done = FALSE)

# "call" our bingo numbers but stop if we have a bingo
for (j in 1:length(bingo_numbers)) {
# for (j in 1:50) {
    my_num = bingo_numbers[j]
    # print(my_num)
    bingo_cards = bingo_cards %>%
        mutate(called = (called | value == my_num))
    
    # check rows for bingo
    do_we_have_rows = check_rows(bingo_cards)
    # adjust record of which cards have bingo
    if (length(do_we_have_rows) == 1) {
        bingo_cards = bingo_cards %>%
            mutate(bingo_done = (bingo_done | do_we_have_rows == card_num))
    } else {
        for (x in do_we_have_rows) {
            bingo_cards = bingo_cards %>%
                mutate(bingo_done = (bingo_done | x == card_num))
        }
    }
    
    # check cols for bingo
    do_we_have_cols = check_cols(bingo_cards)
    
    # adjust record of which cards have bingo
    if (length(do_we_have_cols) == 1) {
        bingo_cards = bingo_cards %>%
            mutate(bingo_done = (bingo_done | do_we_have_cols == card_num))
    } else {
        for (x in do_we_have_cols) {
            bingo_cards = bingo_cards %>%
                mutate(bingo_done = (bingo_done | x == card_num))
        }
    }
    
    if (sum(bingo_cards$bingo_done) == 2500) break
}

# "winning" card numbers that haven't been called yet
# winning for the second part is actually last to win
if (do_we_have_cols) {
    last_card = do_we_have_cols
} else {
    last_card = do_we_have_rows
}

uncalled = bingo_cards %>%
    filter(called == FALSE & card_num == last_card) %>%
    summarise(win_card = sum(value))

# part two answer
uncalled$win_card[[1]] * my_num

```
